Methods with preconditioning with weight decay regularization

### 2023-04-12 paper/main.tex
update paper and code before review

@@

\section{Introduction}
In machine learning we consider unconstrained optimization problem
\begin{equation}
	\min_{w \in \mathbb{R}^d} f(w)
\end{equation}
Problems of the form (1) cover a plethora of applications, including empirical risk minimization,
deep learning \citep{lecun2015deep}, and supervised learning \citep{cunningham2008supervised} tasks such as regularized least squares \citep{rifkin2007notes} or logistic regression \citep{shalev2014understanding}. This minimization problem can be difficult to solve, particularly when the number of training
samples n, or problem dimension d, is large, or if the problem is nonconvex.

Second order preconditioners methods are methods when instead of calculate true hessian of our function for newton method \citep{dennis1977quasi}, some heuristics is used to calculate pseudo-hessian. There are second order preconditioners methods for solving this problem, such as Adam \citep{kingma2014adam} , OASIS \citep{goldberg2011oasis}, AdaHessian \citep{yao2021adahessian}.

Throughout this work we assume that each f $:\mathbb{R}^d \rightarrow \mathbb{R}$ is twice differentiable and also L-smooth. This is formalized in the following assumption.

\begin{assumption}{1}{Convex}
	The function f is convex, i.e. $\forall w, w' \in \mathbb{R}^d$
	\begin{equation}
		f(w) \geq f(w') + \langle \nabla F(w'), w-w' \rangle
	\end{equation}
\end{assumption}

\begin{assumption}{4}{$\mu$ - strongly convex}
	The function f is $\mu$-strongly convex, i.e., there exists a constant $\mu > 0$ such that $\forall w, w' \in \mathbb{R}^d$
	\begin{equation*}
		f(w) \geq f(w') + \langle f(w'), w-w' \rangle +\frac{\mu}{2} ||w - w'||^2
	\end{equation*}

\end{assumption}

@@

\begin{algorithm}[H]
	\caption{Based algorithm}\label{alg:Based}
	$w_0, D_0, \eta_0$
	\begin{algorithmic}
		\For{$k = 1, 2, ...$}
		\State $D_k$ - update by using information of $f(w)$
		\State $\eta_k$ - update
		\State Set $w_{k+1} = w_k - \eta_k D_k^{-1} \nabla f(w_k) - \eta \nabla r(w)$ 
		\State or like that
		\State Set $w_{k+1} = w_k - \eta_k D_k^{-1} (\nabla f(w_k) + \nabla r(w))$
		\EndFor
	\end{algorithmic}
\end{algorithm}

@@

\section{Theory}
To proof this algorigthm we use simplified form of AdamW algorithm
\begin{algorithm}[H]
\caption{AdamW}\label{alg:AdamW}

@@

\end{theorem}
\begin{proof}

>>>>

% Цвета 
\usepackage{xcolor}
\usepackage{color}
\usepackage{colortbl}

@@

\section{Introduction}
In machine learning we consider unconstrained optimization problem
\begin{equation}
	\min_{w \in \mathbb{R}^d} f(w)
\end{equation}
Problems of the form (1) cover a plethora of applications, including empirical risk minimization,
deep learning \citep{lecun2015deep}, and supervised learning \citep{cunningham2008supervised} tasks such as regularized least squares \citep{rifkin2007notes} or logistic regression \citep{shalev2014understanding}.

The classic base method for solving the optimization is gradient descent, but this minimization problem can be difficult to solve, particularly when the number of training samples n, or problem dimension d, is large.
In modern machine learning especially large problems represent the greatest interest.
For such cases stochastic gradient descent \citep{bottou2010large} became popular solution.
Despite its simplicity, it proved itself to be an efficient and effective optimization method.
For a long time first-ordered methods were most popular approach of solving optimization problems.

Other way of solving the problem are methods with adaptive gradient \citep{wilson2017marginal}. These methods posses several superiority over first-ordered methods.
Firstly, they have bigger potential of distributed solving, because first ordered methods spend majority of time on "communication".
Secondly, they are less sensitive to the choice of hyperparameters up to the point that hyperparameters can be set equal to one.
Lastly, this methods often simply show faster convergence on modern large optimization problems, especially this methods became applicable in neural networks solving.
Nowadays it is known that preconditioning methods often outperform other methods on modern large optimization problems \citep{zhang2018three, yao2021adahessian, kingma2014adam, goldberg2011oasis}.

Preconditioning methods refer to techniques that involve scaling the gradient of a problem by a specific matrix $D_t$, which enables the gradient to take into account the geometry of the problem. In the classical case $D_t = (\nabla^2 f)^{-1}$, which corresponds newton's method, however hessian is difficult to calculate and even more difficult to reverse, because of that some heuristics are used to replace the reversed hessian \citep{dennis1977quasi}. In OASIS \citep{goldberg2011oasis} or AdaHessian \citep{yao2021adahessian} hessian is assumed to have diagonal dominance. In Adam \citep{kingma2014adam} gradient is simply normalized, etc. This heuristics were proved to be effective and efficient. General scheme of methods with preconditioning can be framed in the following algorithm

% записать в одну строчку, и пояснить g_t, D_t^{-1}
\begin{algorithm}[H]

            \caption{General scheme for preconditions methods}\label{alg:genalg}

            \begin{algorithmic}
            \small{
            \Require{$\eta, f$}

            \While {$w$ not converged}
            \State $t = t+1$
            \State $g_t \gets \nabla f(w_{t-1})$
            \State $D_t \gets$ pseudo-hessian, precondtioning, scaling
            \State $w_t = w_{t-1} - \eta \cdot g_t D_t^{-1}$    
            \EndWhile
            }
\end{algorithmic}
\end{algorithm}

Regularization is a powerful technique in machine learning that aims to prevent overfitting by adding additional constraints to the model.
It has been widely applied to various machine learning problems, including image classification, speech recognition, and natural language processing, and has shown its effectiveness in improving the generalization capability of neural networks \citep{poggio1987computational}.

In methods with preconditioning appears to be several ways to include regularization.
We can include regularizer $r$ in $g_t$ calculation so it will be taken into consideration while calculating $D_t$. This method is equal to considering optimization problem

\begin{equation}
	\min_{w \in \mathbb{R}^d} f(w) + r(w).
\end{equation}

Or we can include regularizer only on last step, decreasing norm of $w$ \citep{loshchilov2017decoupled}.
This way of regularization is called weight decay and surprisingly turns out to be more efficient in practical problems.
There are few other ways of considering regularizer which will be discussed further in the paper.

In general, our paper provides insight into comparison of different consideration ways of regularization is methods with preconditioning. Here, we provide a brief summary of our main contributions:

\begin{itemize}
    \item \textbf{Proof of preconditioned methods' with weight decay convergence}.  We derive convergence guarantees for preconditioned methods considering assumptions of smoothness, strongly convex and PL-condition.

    \item \textbf{Research of the loss function} Comparison of accuracy and loss function for AdamW and AdamL2. As a result we saw that AdamW asymptotically converges to a non-zero value

    \item \textbf{Competitive Numerical Results} We investigate the empirical performance of Adam's variation including new one on a variety of standard machine learning tasks, including logistic regression.
\end{itemize}

@@

\begin{algorithm}[H]

    \caption{General scheme for preconditions methods}\label{alg:genalg}

    \begin{algorithmic}
    \small{
    \Require{$\eta, \epsilon, f, r$}

    \While {$w$ not converged}
    \State $t = t+1$
    \State $g_t = \nabla f_t(w_{t-1})$
    \State $D_t =$ \textcolor{red}{diag($\sqrt{g_t \odot g_t} + \varepsilon$)}  \hfill \textcolor{red}{AdamW}
    \State $D_t$ = \textcolor{cyan}{$\mathbb{E}[z^T \nabla^2f(w_{t-1}) z]$} \hfill \textcolor{violet}{OASIS}, where $z$ in Rademacher distribution
    \State $w_t = w_{t-1} - \eta \cdot g_t D_t^{-1}$    
    \EndWhile
    }
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]

    \caption{Adam($\lambda$)}\label{alg:Adam}

    \begin{algorithmic}
    \small{
    \Require{$\eta, \beta_1, \beta_2, \epsilon, f, r$}
    %\State $m_0 = 0$ -- 1-st moment vector
    %\State$v_0 = 0$ -- 2-nd moment vector
    \While {$\theta$ not converged}
    \State $t = t+1$
    \State $g_t = \nabla f(w_{t-1})$ + \textcolor{blue}{$\nabla r(w_{t-1})$} \hfill \textcolor{blue}{AdamL2}    
    \State $m_t = \beta_1 \cdot m_{t-1} + (1 - \beta_1) \cdot g_t$
    \State $v_t = \beta_2 \cdot v_{t-1} + (1 - \beta_2) \cdot g_t^2$
    \State $\hat{m_t} = \frac{m_t}{1-\beta_1^t}$ + \textcolor{green}{$\nabla r(w_{t-1})$} \hfill \textcolor{green}{MyAdamW}
    \State $\hat{v_t} = \frac{v_t}{1-\beta_2^t}$ 
    \State $w_t = w_{t-1} - \eta \cdot \frac{\hat{m_t}}{\sqrt{v_t} + \epsilon}$    - \textcolor{red}{$\eta \nabla r(w_{t-1})$ } \hfill \textcolor{red}{AdamW}
    \EndWhile
    }
\end{algorithmic}
\end{algorithm}

@@

\section{Theory}
\begin{assumption}{1}{Convex}
	The function f is convex, i.e. $\forall w, w' \in \mathbb{R}^d$
	\begin{equation}
		f(w) \geq f(w') + \langle \nabla F(w'), w-w' \rangle
	\end{equation}
\end{assumption}

\begin{assumption}{4}{$\mu$ - strongly convex}
	The function f is $\mu$-strongly convex, i.e., there exists a constant $\mu > 0$ such that $\forall w, w' \in \mathbb{R}^d$
	\begin{equation*}
		f(w) \geq f(w') + \langle f(w'), w-w' \rangle +\frac{\mu}{2} ||w - w'||^2
	\end{equation*}
\end{assumption}

To proof this algorigthm we use simplified form of AdamW algorithm
\begin{algorithm}[H]
\caption{AdamW}\label{alg:AdamW}

@@

\end{theorem}

\begin{theorem}
    Suppose the Assumption 1, 2, 4, 5 and let $\varepsilon > 0$ and let the step-size satisfy
    \begin{equation*}
        \eta \leq \frac{2 \alpha}{\tilde{L}}
    \end{equation*}
    Then, the number of iterations performed by AdamW algorithm, starting from an initial point $w_0 \in \mathbb{R}^d$ with $\Delta_0 = \tilde{F}(w_0) - \tilde{F}^*$, required to obtain and $\varepsilon$-approximate solution of the convex problem (link here to problem 1) can be bounded by
    \begin{equation*}
        T =  \mathcal{O}\left( \frac{\ln \frac{\Delta_0}{\epsilon}}{2 \mu \eta^2(\frac{1}{\eta} - \frac{\tilde{L}}{2 \alpha})} \right)
    \end{equation*}
\end{theorem}
\section{Appendix}

\begin{theorem}
Suppose the Assumption 1, 2, 5 and let $\varepsilon > 0$ and let the step-size satisfy
\begin{equation*}
    \eta < \frac{2 \alpha}{L + l \cdot \alpha} 
\end{equation*}
Then, the number of iterations performed by AdamW algorithm, starting from an initial point $w_0 \in \mathbb{R}^d$ with $\Delta_0 = \tilde{F}(w_0) - \tilde{F}^*$, required to obtain and $\varepsilon$-approximate solution of the convex problem (link here to problem 1) can be bounded by
\begin{equation*}
      T = \mathcal{O}\left( \frac{2\Delta_0 \Gamma \alpha } {(2\alpha - \tilde{L}\eta) \eta \varepsilon} \right)
\end{equation*}

\end{theorem}

\begin{proof}
